---
title: "An Amalgamation of chpts 2-4!"
author: "Rae Dunbar"
date: "10/29/2022"
output: html_document
---

In this data frame we will use programming and techniques learned in chapters two to four. 

## Setting up R
```{r}
rm(list=ls()) #clear R's memory
#Adding necessary packages
library(here)
library(dplyr)
library(ggplot2)
```

## Pulling in the Data

Given this is due on Halloween we will be working with Halloween data! 

```{r}
here() #See where R is looking
halloween<- read.csv(here("Problem-Set-2", "Data for the code", "Halloween_data_PS2.csv"), stringsAsFactors = TRUE) 
#I needed to first bring R into my Problem set 2 folder, since R was looking just in the project. 
#Given we wanted both Costume and Type as factors and not as character, I did stringsAsFactors=TRUE to make them factor variables right away.
```

## Examine the Data

```{r}
names(halloween) #shows column names
```

All column names seem to spelled correctly, and there are no dots indicating unwanted spaces, so we can keep the column names. 

It looks like the data has four variables, the costume, Type, the number of candy pieces, and the age of what I presume is the costume wearer. To understand what "Type" means and see the spread of the data lets use summary:

```{r}
summary(halloween) # provides a summary of the data
```
 Type appears to be whether the costume is scary or not. There are four levels for 'Type' and three of which are 'scary.' I see the error in one of the entries, since it has the S capitalized, but the other two look exactly the same. Lets see if running levels helps clarify that. 
```{r}
levels(halloween$Type) #shows the choices available for the column.
```

Now I see there is a space right after one of the 'scary' datum. So, we need to fix that to have the correct number of levels in that column. For the other two variables, the max number of candy peices being 16 is probably an outlier, given the mean is 2.9 and the third quartile is only 4. The age data look reasonable. Lets double check the possbility of an outlier by looking at more of the data with another.

```{r}
tbl_df(halloween) # make a table of your data
```
Aside from the wopping 16 candies, the biggest number is 6, making 16 an outlier. The age distribution is, as I aspected based on the mean, reasonable.There are 13 and 14 year old children, with the eldest being 15. It is feasible one child was particularly victorious with candy,  so we will leave the outlier in the data. Based on how it impacts the figures we may want to leave it out later. 

I noticed another issue however, some entries are spelled differently, so there are costumes that are the same, but R thinks they are different. 

## Fixing the Data

the only issue that requires attention is fixing the factors in the Type column.
```{r}
sfix<-which(halloween$Type== "Scary" | halloween$Type== "scary " )
halloween$Type[sfix]<- "scary"
```

### checking that it was fixed

```{r}
tbl_df(halloween)
```
It looks like it worked! So, now we need to re-asign the factors so R can keep up with our changes. 

### Change the Factors 
```{r}
halloween$Type<- factor(halloween$Type)
# Assign the factors that are now in the column Type to the column Type. 
```

```{r}
levels(halloween$Type)
```

### Fixing the Costume types


```{r}
levels(halloween$Costume)
```
It looks like witch, creepy clown, and wonder woman need to be fixed. I am not sure if the difference between clown and creepy clown should be left or not. Looking at the last figure, there is a creepy clown category, but not a clown category. So I will change them all to Creepy clown. I will base which version of the costume types I choose off of the last figure that we need to replicate. 

```{r}
cfix<-which(halloween$Costume== "creepy clown" | halloween$Costume== "clown" )
halloween$Costume[cfix]<- "Creepy clown"

wfix<-which(halloween$Costume== "witch" | halloween$Costume== "witch " )
halloween$Costume[wfix]<- "Witch"

wwfix<-which(halloween$Costume== "wonder woman" | halloween$Costume== "Wonder Woman" )
halloween$Costume[wwfix]<- "Wonder woman"
```

### Reseting the Levels

```{r}
halloween$Costume<- factor(halloween$Costume)
levels(halloween$Costume)
```

Now we have the proper number of factors


## Calculating Mean
```{r}
halloween_means<-halloween %>% group_by(Costume, Type) %>% summarize(mean_candy=mean(candy_pieces, na.rm=TRUE), mean_age=mean(age, na.rm=TRUE))
#na.rm tells R to ignore the NA data and calculate the mean anyway.
```
The zombie costume had the highest mean candy count taken at 6.86, though the outlier may have skewed that. And the unicorn costumes grabbed the least amount at 1.17. Creepy clowns had the highest mean age at 13.2 with rabbits having the lowest, 5.0 mean age. Lets see this in graph form. 

## Making Graphs

### Graph 1
Use the mean data to make the first graph. You could combine the code from above (halloween %>% group_by(Costume, Type) %>% summarize(mean_candy=mean(candy_pieces, na.rm=TRUE), mean_age=mean(age, na.rm=TRUE))) into the data= portion of the code below, but then you would not be able to see the means in a another data frame, which I interpreted you wanted, so I left these as two separate code chunks.
```{r}
ggplot(data=halloween_means, aes(x=mean_age, y=mean_candy, colour=Type)) + xlab("Mean age of kid (years)") + ylab("Mean number pieces taken")+ geom_point(size=3) + theme_bw()



#use ggplot to layer a graph together first taking the data from the halloween_means table, make the x and y axis as noted and call them what is noted with xlab and ylab.Then add points, making the size 3, which seems the closest to the size you chose, and the theme of the graph black and white colored.  
```
Figure 1. Mean number of pieces of candy taken as a function of age of trick-or-treater and costume type (Red dots = costumesthat are not scary; blue dots = scary costumes).

There is a weak correlation between mean age of kids and number of candy they take, with older kids taking more candy. 

### Graph 2
Our task is to write code that isolates the data for children in “not scary” costumes who are 10years old or younger and use the resulting data to plot a histogram showing the number of kids in each age class and that comes as close as you can to replicating figure 2, below.

We need to work with the origonal halloween data for this.
```{r}
histogram<-halloween %>% filter(halloween$Type=="Not scary" & halloween$age<=10)
# take the halloween data and show only the data where the type is not scary and the child is 1o years or younger in a new dataframe called histogram. 
```

Now we can plot this: 

```{r}
ggplot(data=histogram, aes(x=age)) + xlab("Age of child") + ylab("Number") + geom_histogram(binwidth=2)
```
Figure 2. Number of children age 10 and under wearing costumes that are not scary. 

The peak of the histogram is at age 6. So most of the children wearing non scary costumes are around six years old, with fewer older children wearing scary costumes. This makes sense, as older children probably want to be scary at Halloween.

The only feature I am missing on this histogram is the lines separating the different bars. 

### Graph 3

write code that replicates figure 3 below as closely as you can (Hint: try `coord_flip()` in your ggplot to get the axes reversed so that you can read the costume choice names). 
```{r}
ggplot(data=halloween, aes(x=Costume, y=age)) + ylab("Child age(years)") + xlab("Costume choice") + geom_boxplot() + geom_point(size=4, colour='lightgrey', alpha=0.6) + facet_wrap(~Type)+ coord_flip()
#facet_wrap splits it into two figures side by side
#geom_point() adds points to the data with the color being light grey and slightly see through using alpha. 
#coord_fip switches around the x and y axis. Another option, is to just make x=age and y=Costume. 
```
Figure 3. Box and whisker plot showing the age distribution of children wearing different Halloween costume choices, separated according to whether the costume is scary or not. 

We can see based on this figure that older children, approx. above 6 years old, tended to wear scary costumes, with the exception of one witch datum, and the wonder women data. Though, wonder women can be scary, given her strength and all. The correlation is again weak, a large sample size is needed. The box plot and points allow us to see the spread of the data. For instance, children dressed as zombies varied from 8-15, wherease creepy clown dressed children has a smaller spread of 12-14. 
